Parental Migration and Left-Behind Children's Cognitive and Academic Performances

### 2022-07-07 MScThesisText.Rmd

---
title             : "Left-Behind Children's Educational Performance Drops Slightly After Parents Migrated"
shorttitle        : "Left-Behind Children"

@@

abstract: |
  Millions of children worldwide are "left behind" when their parents migrate away for work. 
  Parents' labor migration increases household income but decreases parental care, thereby exerting mixed influences on child development, including educational performance. 
  To what extent do different arrangements of parental migration affect children's schooling performance in the short term? 
  Studying the case of China, I obtained a sample of four thousand junior high school students from a nationally representative two-wave panel survey. 
  I differentiated household migration arrangements of any one parent, of mother only, of father only, and of both parents.
  I measured children's educational performance by standardized cognitive test score and academic exam grades.
  For causal inference, I built difference-in-differences models with propensity score matching. 
  Results show that the new left-behind children's performance at school dropped slightly compared to that of non-left-behind peers. 
  Father-only migration is negatively associated with children’s cognitive abilities, while mother-only migration is negatively associated with children’s academic grades.

@@

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("cobalt")
r_refs("references.bib")
```

@@

Worldwide, millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work. These are the so-called "left-behind children" or "stay-behind children." They are common in migrant-sending regions, such as Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. Researchers estimated that one-sixth of Mexican children [@dewaard2018], one-fifth of Bulgarian children [@popova2018], and one-fourth of Chinese children were living apart from their migrating parents [@nbs2017]. In a 2021 resolution, the Council of Europe's @parliamentaryassembly2021a expressed concern about "the scale of this phenomenon and the long-term damage it creates" (p. 2) and declared that leaving "millions of labour migrants' children without parental care is a mass violation of human rights" (p. 1).

How are stay-behind children faring when parents have migrated? Research results are mixed due to the differential effects of parental migration. Consider children's education. According to the review articles by @debrauw2019 and @vanhook2020, labor migrants send remittances back, which can fund children's schooling as well as improve their nutritional and living conditions. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration decreases the amount and quality of parental guidance, protection, and support. It is uncertain how these two opposing effects together shape children's educational performance, which merits empirical investigation.

A challenge arises in assessing the causality between parent's migration and children's educational achievement. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, a quasi-experimental design can help with causal inference. The cause should be on par with random assignment to participants in the analysis, conditional on identification assumptions [@cunningham2021].

In this thesis, my research question is: To what extent do different arrangements of parental migration affect the educational performance of left-behind children in the short term? I choose China as the case and analyzed data from a two-wave panel survey. The nationally representative sample includes children in both urban and rural areas. I develop a difference-in-differences model for causal inference and applied propensity score matching for robustness check. The outcome was measured with children's cognitive test score and academic exam scores. I demonstrate that parental migration reduces the educational performance of children staying behind, but only to a small extent.

@@

## Theoretical Framework

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One is the resource generation model. It holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic capital is transferable to human capital, which is the stock of "skills, talents, health, and expertise" in people that enhance their productivity [@goldin2016; @botev2019]. @yang2008 noted that in the Philippes, a sharp depreciation of peso increased the amount of remittances from migrant laborers working abroad. This subsequently boosted children's school enrollment and educational expenditure. In China, @chang2019 interviewed migrant workers and their family members. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p.1]. Parental migration decreases nurturing care, as characterized by less homework supervision and emotional support by parents [@wen2019]. Hong talked to a group of left-behind children at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to grades and disengaged with school. Some even longed to quit school to pursue a free, independent life (p. 14).

Given the two opposing effects, researchers have obtained mixed results regarding how parental migration impacts children's education. @botezat2020 studied the case of Romania and found a positive effect of parents' migration on academic grades of adolescents aged eleven to fifteen. @wang2019 collected data from junior high school students in central China and found no effect of parental on left-behind children's achievement in mathematical test. Studying four Chinese provinces, @debrauw2017 showed that as ample migration opportunities appeared in villages, enrollment to senior high school decreased. @nguyen2016 demonstrated that parental migration was associated with lower cognitive capabilities among Indian and Vietnamese children aged five to eight. The inconsistency can be complicated by factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

Parental migration immediately disrupts family structure, which in turn lowers academic performance of children. It takes longer time for parental migration to generate resources that improve children's educational attainment. I propose the first hypothesis:

$H_1$ In the short term, parental migration negatively affect the cognitive development and academic performance of children left-behind.

Researchers have studied whether the impact differ between mother absence and father absence induced by labor migration. @xu2019 adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). @chen2019 conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

Studies noted that in some Asian and Latin American societies, the social norms designate men as breadwinners and women as caregivers. Under such norms, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left-behind children may even resent their migrant mothers for transgressing on the expected care-giving roles. I propose the second hypothesis:

$H_2$ In the short term, mother's migration negatively affect the cognitive development and academic performance of children left-behind, while father's migration does not show a effect.

## China's Context

Three features make China a suitable case for studying left-behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers [@nbs2022]. They form the backbone of essential industries, such as retail, delivery, transportation, and manufacturing. The number of left-behind Chinese children reached 68.7 million in 2015 [@nbs2017]. The internal migration stemmed from the neoliberal economic regime.

The neoliberal economic regime sounds at odds with China, a so-called socialist country. Actually, in the 1980s, the ruling communist party ditched the planned economy, initiated market reforms, and opened up the country. Since then, China has integrated in the global economy and contributed substantial growth to it. The economic reform and globalization have enriched the coastal cities like Shanghai and Shenzhen. The rural hinterlands, however, have lagged behind. The regional inequality needs social policy to amend, but China's welfare programs are discriminatory -- it segregates the population. Those who reap the greatest benefits are public servants, urban dwellers, and formally employed people. As rural people migrate to the city, they find themselves with limited access to social welfare in the destination. The municipal governments only provide welfare to local residents and expect non-local residents access public services elsewhere [@duckett2020; @zhang2018].

The institutionalized discrimination against migrant laborers find its roots in China's "multilevel citizenship" system. As @vortherms2021 demonstrated, the constituents of citizenship -- "membership rules and rights entitlements" -- operate below the national level in China. The local state confers most of the rights and defines who is eligible to these rights through the *hukou*, or household registration system:

> A political institution rooted in centuries of Chinese bureaucracy, the modern hukou is the primary identity document defining membership and entitlements for local citizenship. Accessing fundamental citizenship rights depends on your registration status ... Registration includes basic identifying information, a specific address of registration, and a type: urban, rural, or resident. [@vortherms2021]

In other words, the hukou system binds one's citizenship rights with one's place of registry. In theory, anyone can move out of town freely. In practice, they cannot change hukou freely. When rural residents move to the city, it takes months -- if not years -- to apply for a new hukou. Before receiving urban hukou, they can hardly send their children into public schools in the host city [@gu2021]. Migrant workers had organized petitions, strikes, and protests, only to fall on deaf ears or even end with police crackdown [@chan2017]. Facing financial and policy constraints, many migrant laborers have to leave children in their hometowns.

The existing research about China's left-behind children has three major shortcomings. The first is that most literature focuses on children left behind in rural areas while neglecting the urban counterparts. The National Bureau of Statistics stated that the number of urban left-behind children had been growing and expected that the trend would continue [@nbs2017]. My sample includes urban left-behind children as well. The second shortcoming lies in sampling and research design. Some studies suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one to a few towns [@liu2021; @yang2022; @shu2021]. Others only estimated correlation using cross-sectional data [@akezhuoli2022; @jin2020]. These studies could have employed causal inference techniques like fixed effect, matching, or inverse probability of treatment weighting. The third shortcoming lies in the measurement of outcome. @chang2019 and @wang2019 measured students' educational performance with the test score of only one academic subject, namely mathematics. To tackle these limitations, I adopted a nationally representative two-wave panel dataset. I included the test results of multiple subjects to measure comprehensive academic achievement.

@@

## Sampling

I obtained the data from the China Education Panel Survey (CEPS) [@nsrc]. It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The survey team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. The follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed `r printnum(nrow(d))` of the seventh grade cohort. I further restricted the sample to students staying with both parents at the original domicile at the baseline survey. To deal with missing values, I adopted list-wise deletion for dependent, independent, and control variables. The final sample size amounted to `r printnum(nrow(d_cl))` .

@@

I explored two categories of educational performance: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on the Item Response Theory and then standardized the score across the whole national sample with a mean of zero and standard deviation of one [@wang2015].

Another dependent variable concerned students' academic performance. This was measured by the total scores of mid-term exams in Chinese, mathematics, and English as provided by each school. Exam scores are reliable indicators of learning performance, and are comparable across schools for the same cohort of students. This is related to how junior high schools operate as part of the compulsory education system in China. According to @oecd2016, China's junior secondary education is a unified system that does not distinguish between academic and vocational tracks. At school, Chinese, mathematics, and English are core subjects and others as supplementary subjects. Throughout junior high school, students study the core subjects in preparation for the senior high school entrance exam. Within each cohort in one particular school, teachers use the same syllabus and administer the same exams during a given assessment period. Across cities and regions, junior high schools adopt the educational content and assessment set by the Ministry of Education [@oecd2016].

I transformed the exam grades through the following steps. I first aligned different grading scales (full mark of 100, 120, and 150 for one subject) to the hundred-point scale. After the conversion, I checked for abnormal values outside the 0-100 range. `r printnum(num.outlier$n)` observations were outliers, accounting for less than one percent of the sample. I trimmed the values of subject exam score with minimum of zero and maximum of one hundred. In this way, I can limit the impact of outliers while retaining these observations in the sample. After trimming, I aggregated the subject exam scores to obtain a total score of a 300-point scale.

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed four treatment dummy variables measuring migration arrangements: any parents absent, only father absent, only mother absent, and both parents absent. These arrangements are mutually exclusive. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

The control variables on children include their age, gender, ethnicity, self-rated health, hukou type, number of siblings, whether the child had skipped grade or repeated grade, and the level of education that the child aspires to achieve. At the household level, I controlled the level of education of the best-educated parent, the parental expectation on the child's level of education, the perceived household financial conditions, the number of extracurricular books at home, and the internet access at home.

@@

I adopted `R` for all the analyses, the `fixest` package for econometric modeling [@berge2022], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i,j;1} + \lambda \cdot C_{j} + \theta \cdot C_{k} + \varepsilon_{i,j,k}$$ For student $i$ in class $j$ in baseline and class $k$ in endline, $\Delta score_{i,j}$ is the change in test score between baseline and follow-up surveys, $migr_{i}$ is the treatment dummy variable, $score_{i,j,k;1}$ is the test score at baseline, and $C_{j}$ and $C_{k}$ are the classroom fixed-effects at baseline and follow-up, respectively. Classroom fixed-effect can account for differences due to time-invariant factors at the classroom level and above, as students nested in the same class are also in the same school and county. About one-sixth of the students were reassigned into new classroom at the end of baseline academic year. This necessitates adding the wave-two classroom fixed-effect.

The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein2022]. The standard errors were clustered at the classroom level.

In addition, I present an unrestricted and adjusted DID model: $$\Delta score_{i,j,k} = \alpha + \beta \cdot migr_{i} + \gamma \cdot score_{i,j,k;1} + \lambda \cdot C_{j} + \theta \cdot C_{k} + \zeta \cdot X_{i} + \varepsilon_{i,j,k}$$ where $X_{i}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

Table 1 below summarizes the changes in migration status in the sample. All `r printnum(nrow(d_cl))` children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

```{r i-var}
table_independent_variable
```

: Parental migration status at follow-up survey.

Tables 2 and 3 below provide the whole-sample summary statistics. In addition, the violin plots in the appendix visualize the distribution of outcome variables for the treatment group and the control group.

```{r d-var}
table_dependent_variable
```

: The independent variables

```{r c-var}
table_control_variable
```

: The control variables

## Regression Analysis

The unrestricted DID models support my hypothesis one -- newly left-behind adolescents are more likely to perform worse in cognitive and academic abilities than their non left-behind counterparts. If a household sees any parent migrated, the child's cognitive test score and total exam grade would decrease relative to that of children whose parents stay with them, holding everything else constant. Parental migration is linked to `r round(coef(did_cog)[1],2)` SD in cognitive test score and `r round(coef(did_cog)[1],2)` points in academic exam grades. Both relationships are slight in strength.

The results only partially support the hypothesis two -- mother migration has a more negative effect than father migration. This holds true for academic abilities. Children from only mother migrate households are associated with `r round(coef(did_cog_m_only)[1],2)` points relative to that of children whose parents stay along. The performances of children with both parents migrated or only father migrated do not show statistically significant difference from the non left-behind children. Regarding cognitive abilities, the reverse is true. Children with only mother migrated do not show statistically significant difference from the non left-behind children. Children with only father absent are associated with (`r round(coef(did_cog_f_only)[1],2)` SD), and both parents absent (`r round(coef(did_cog_both)[1],2)` SD) compared to the non left-behind children.

```{r reg-cog}
table_cog
```

```{r reg-ttl}
table_ttl
```

The adjusted DID models produce similar results compared to the unadjusted models, with only one exception. That is, for the cognitive skills, the both-parent-absent coefficient saw a change of statistical significance level from five percent level to ten percent level.

Some predictors are strong in both models of cognitive and academic abilities. These are children's age and educational aspiration as well as parents' expectation on children. In terms of age, the older students' scores drop relatively more than younger students'. One year older in age is associated with `r round(coef(did_adj_cog)[3],2)` SD for standardized cognitive test and `r round(coef(did_adj_ttl)[3],2)` points for academic exams. Parent's expectation for children completing university and children's own aspiration for finishing university are associated with higher scores. The effect sizes are `r round(coef(did_adj_cog)[15],2)` SD and `r round(coef(did_adj_cog)[12],2)` SD for standardized cognitive test, respectively; and `r round(coef(did_adj_ttl)[15],2)` points and `r round(coef(did_adj_ttl)[12],2)` points for academic exams, respectively.

Several predictors are strong in the model of academic grades but not in that of cognitive scores. These are children's gender and self-reported health, and the years of parental education. Girls outperform boys by `r round(coef(did_adj_ttl)[4],2)` points in academic exams, but do not show advantage in cognitive test. Students who consider themselves healthier are associated with higher scores in academic exams. Parents with more education are associated with their children's higher scores in academic exams. If the years of parental education of the best-educated among the two increase by one, it would be linked to a `r round(coef(did_adj_ttl)[14],2)` points increase in the child's academic exam grade. These three predictors are not significant for cognitive outcomes.

Preschool attendance is associated with higher cognitive outcomes `r round(coef(did_adj_cog)[8],2)` SD, but its effect for academic outcomes is not significant.

```{r reg-cog-adj}
table_adj_cog
```

```{r reg-ttl-adj}
table_adj_ttl
```

## Matching

Parental migration is not a randomly assigned treatment. A possible circumstance is that smart and ambitious students perform better at school; their high performance and ambition prompt parents to migrate to support children's schooling. Another scenario is that poverty may hamper children's educational achievement while driving parents to migrate. To tackle the selection bias herein, an applicable technique is matching. This method makes two groups comparable by pairing a unit in the treatment group with one or several units in the control group based on observable covariates.

Popular choices for matching include coarsened exact matching (CEM) [@waldman2022] and nearest-neighbor propensity score matching (PSM) [@xu2015]. As @greifer2022 introduced, CEM is superior in achieving balance as it optimizes the entire joint distribution of covariates. The downside is that it leaves many units without an exact match and discards these unmatched units. The nearest-neighbor matching examines all the treated units and selects the closest control unit to be paired. The eligibility as the "closest" can be determined with the differences in propensity score. Propensity score denotes the probability of receiving treatment based on the specified covariates. Nearest-neighbor PSM does not try to optimize any condition; each pairing ignores how other pairing will occur or have occurred [@greifer2022]. @king2019 argued against this matching method as it risks more "imbalance, model dependence, and bias" for highly balanced data.

>>>>

---
title             : "Parental Migration and Left-Behind Children's Cognitive and Academic Performances"
shorttitle        : "Left-Behind Children"

@@

abstract: |

@@

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

@@

```{r setup, include = FALSE}
load('.Rdata')
library("papaja")
library("magrittr")
r_refs("references.bib")
```

@@

Millions of children are resided in their home communities but separated from their parents as either one or both parents have migrated for work [@antia2020]. These are the so-called "left behind children" who are common in migrant-sending regions -- Latin America, Sub-Sahara Africa, East Europe, and large parts of Asia. We do not know the global size of left behind children population, but we can find estimate from in-country surveys. [@dewaard2018] analyzed the Mexican and Latin American Migration Projects and estimated that around fifteen percent of children were left behind in Mexico, El Salvador, Nicaragua, and Puerto Rico in the 2000s. [@popova2018] cited a 2016 study by UN Children's Fund Bulgaria indicating that one-fifth of Bulgarian children had one or both parents working abroad. In China, the 2015 one-percent National Population Sample Survey showed that one out of four children lived under parental absence induced by migration [@nbs2017].

The issue of left behind children has attracted public attention. In March 2021, the Parliamentary Assembly of Council of Europe [-@parliamentaryassembly2021] passed a resolution that

> deplores the complacency of both countries of origin and countries of destination of labour migration, which tend to tolerate excessive labour migration as it has significant short-term benefits in the form of remittances for the former, and of a cheap, flexible labour force for the latter. This situation is not acceptable, and it is not sustainable. Leaving millions of children without parental care is a mass violation of human rights and an unnecessary threat to the stability and prosperity of our countries. (p. 1)

The resolution suggests that parental migration impacts child development by exerting different effects that work oppositely. Consider children's education. According to Van Hook and Glick [-@vanhook2020], labor migrants send remittances back home, boosting the family income. Higher income can fund children's schooling as well as improve their nutritional and living conditions. All these may enhance children's educational outcome. The inflow of remittances, however, cannot substitute for the loss of parental care. Parental migration halts the day-to-day, face-to-face contact between parents and children. It decreases the amount and quality of parental guidance, protection, and support. All these may hinder children's school performance. It is uncertain how these two opposing effects shape children's educational performance, which merits empirical investigation.

A challenge arises in when we want to assess the causality between parental migration and children's education. Experiments can unveil causal relations, but it is unfeasible and unethical to design an experiment that randomly assigns parents to migrate without their children. Alternatively, we can infer causality from a quasi-experimental design -- that is, when the cause is on par with random assignment to participants in the analysis, conditional on identification assumptions [@cunningham2021].

Three features make China a suitable case for studying left behind children. The first feature is the huge scale of internal labor migration. According to the National Bureau of Statistics, the year 2021 recorded 292.5 million in-country migrant workers [@nbs2022]. The second feature is the neoliberal welfare regime. Neoliberalism extends the logic of the market to most spheres of human life. Driven by the neoliberal doctrine, the government shifts its responsibilities of social services and welfare to individuals and communities. Since China initiated pro-market reforms in the 1980s, the state has reduced its welfare provisions and passed them to individuals and families [@zhan2020]. The third feature is the authoritarian control against citizens. According to Gu [-@gu2021], the hukou (household registration) system controls people's mobility by binding a household's public welfare access with its place of registry. For a rural household member moving to the city, it takes much time and effort to apply for urban hukou status. Without an urban hukou they cannot send children to public school in the host city. Facing financial and policy constraints, many migrant laborers have to leave children in their original domiciles. The number of left behind Chinese children reached 68.7 million in 2015 [@nbs2017].

To what extent do parental migration affect the cognitive development and academic performance of left-behind children in the short term? I analyze the data from a nationally representative two-wave panel survey. The sample includes children in both urban and rural areas. I develop a difference-in-differences design for causal inference. The dependent variables include children's cognitive test score and Chinese, mathematics, and English exam scores.

@@

Researchers obtained mixed results regarding how migration-induced parental absence affects children's education. Using data collected from junior high school students, Li and colleagues [-@li2017] found that parental migration negatively impacts left-behind children's achievement in mathematical test. Wang and others [-@wang2019] also collected data from students of junior high school, but discovered no overall effect of parental migration on the mathematical test grades of children left behind. Bai and coauthors [-@bai2018] gathered data from primary school pupils. They reported that parental migration "does not appear to have a negative effect" on the English test results of left-behind children, and "in some cases even appears to have a positive effect" (p. 1165). The inconsistency can result from factors such as the left-behind children's age and gender, the migrant parent's original domicile and eventual destination, and the arrangement of which parent migrate for how long.

Between parental migration and children's education, the mechanisms trace to two theoretical frameworks. One model, the resource generation model, holds that parental migration brings forth greater resources in the form of economic capital. Economic capital is, in Bourdieu's words, "immediately and directly convertible into money and may be institutionalized in the form of property rights" [-@bourdieu2002, p.16]. Economic capital is transferable to human capital, which is the stock of "skills, talents, health, and expertise" in people that enhance their productivity [@goldin2016; @botev2019]. [@chang2019] provided evidence through interviews with migrant workers and their family members. The informants agreed that the city offered more jobs and higher pay than did the countryside. Oftentimes, they found that labor migration presented the only way to sustain their children's education and livelihood.

The other theoretical model is the family disruption model. Children flourish under parental nurturing care. This term refers to the "stable environments that promote health and adequate nutrition, protect from threats, and provide opportunities for learning and responsive, emotionally supportive and developmentally enriching relationships" [@black2021, p.1]. Parental migration decreases nurturing care, thereby hindering children's developmental outcomes. The loss of nurturing care slows children's accumulation of human capital. Hong talked to a group of left behind children in grade nine when she conducted ethnography at a rural school [@hong2019]. These students expressed the utter sense of loneliness "with no advocate, no support and with no one looking out for them" (p. 13). Many of them felt indifferent to school grades, and some even longed to ditch school to pursue a free, independent life (p. 14).

I hypothesized that (H1) in the short term, parental migration negatively affect the cognitive development and academic performance of children left behind. This is because parental migration immediately disrupts family structure, resulting in lower academic performance of children. It takes longer time for parental migration to generate resources that improve children's performance at school.

Researchers have studied whether the impact differ between mother absence and father absence in the context of labor migration. Xu and others [-@xu2019] adopted a fixed-effects propensity score weighting model on cross-sectional data collected from seventh and ninth graders in rural areas across China. They found that the absence of father only or both parents had "little or no association with negative outcomes" on children's academic and cognitive performances. Only-mother absence, however, showed a "strong association with negative outcomes" (p. 1646). Chen and others [-@chen2019] conducted structural equation modeling on the cross-sectional data collected from fourth to seventh graders in the countryside of one province. They found that mother migration was "negatively associated with children's social competence and academic performance," while father migration had no direct effects on the two outcomes (pp. 860-861).

I hypothesized that (H2) in the short term, mother migration negatively affect the cognitive development and academic performance of children left behind, while father migration does not show a effect. This relates to the social norms that designate men as breadwinners and women as caregivers. Studies noted that in some Asian and Latin American societies, children are more likely to accept fathers' migration than mothers' [@murphy2022]. Left behind children may even resent their migrant mothers, viewing them as transgressing on the expected care-giving roles.

The research on China's left behind children has three major shortcomings. First, most literature focuses on children left behind in rural areas while neglecting the urban counterparts. [@nbs2017] stated that the number of urban left behind children had been growing and expected that the trend would continue. The second shortcoming lies in data and sampling. Some studies draw from cross-sectional data, making it hard to infer causal relations [@zhou2014; @jin2020]. Others suffer from a limited sample size of less than one thousand participants or a limited geographic coverage of one or a few towns [@liu2021; @yang2022; @shu2021; @jin2020]. The third shortcoming lies in the selection and measurement of outcome. [@chang2019; @wang2019] measured students' educational performance with the test score of only one academic subject, namely mathematics. The test results of multiple subjects should be included to measure comprehensive academic achievement.

@@

## Participants

I obtained the data from the China Education Panel Survey (CEPS) [@nsrc]. It is a nationally representative, school-based survey featuring junior high school students. The survey administrator is the National Survey Research Center at Renmin University of China. The research team adopted multi-stage stratified probability proportional to size sampling and administered a paper-and-pencil questionnaire to each participant. The baseline survey took place in the 2013-2014 academic year with a sample size of about twenty thousand students in the seventh and ninth grades. These students were nested in 438 classrooms of 112 schools in 28 urban districts or rural counties. The follow-up survey in the 2014-2015 academic year lost track of the ninth grade cohort and only surveyed `r printnum(nrow(d))` of the seventh grade cohort. I further restricted the sample to students staying with both parents at the original domicile at the baseline survey. To deal with missing values, I adopted list-wise deletion for dependent, independent, and control variables. The final sample size amounted to `r printnum(nrow(d_cl))` .

@@

I explored two categories of adolescence development: cognitive and academic abilities. One key dependent variable, the student's cognitive skill, was measured by a fifteen-minute standardized test. The test covered three types of reasoning: verbal, visuospatial, and numerical. The CEPS research team calculated the raw score based on Item Response Theory and then standardized the score with a mean of zero and standard deviation of one.

The other set of dependent variables concerned students' academic performance. This was measured by the total scores of mid-term exams in Chinese, mathematics, and English as provided by each school. I aligned different grading scales (full mark of 100, 120, and 150 for one subject) to the hundred-point scale. Accordingly, the total score of three subjects has a 300-point scale.

The key independent variable, the parental migration status, was identified by two items in the parent survey. In both the baseline and the follow-up surveys, parents specified the family members living in the same household at the time. I constructed six treatment dummy variables measuring migration arrangements: both parents absent, only father absent, only mother absent, father absent (unconditional on mother's migration status), mother absent (unconditional on father's migration status). These arrangements can overlap with each other. The control group consisted of parents staying with their children throughout at both the baseline and the follow-up surveys. I then removed the observations that reported parental divorce or death. These type of parental absence are irrelevant to migration.

@@

I adopted `r cite_r("r-references.bib")` for all the analyses, the `fixest` package for building the econometric models [@bergé2018], and the `modelsummary` package for presenting the results [@arel-bundock2022]. I built a two-way fixed effect model in a difference-in-differences (DID) design, following Bai and colleagues' [-@bai2018; -@bai2020] approach. The model served to analyze the educational outcome of children newly left behind by parents vis-a-vis that of children staying together with parents. I first specified an unrestricted and unadjusted model: $$\Delta score_{i,c} = \alpha + \beta \cdot migr_{i,c} + \gamma \cdot FE_{c} + \lambda \cdot score_{i,c;base} + \varepsilon_{i,c}$$ For student $i$ in classroom $c$, $\Delta score_{i,c}$ is the change in test score between baseline and follow-up surveys, $migr_{i,c}$ is the treatment dummy variable, $FE_{c}$ is the classroom-level fixed effect, and $score_{i,c;base}$ is the test score at baseline. The model is unrestricted because it does not restrict on the coefficient associated with the baseline scores. The model is unadjusted as it does not adjust for additional covariates. Theoretically, it is unnecessary to include covariates that vary over group but remain constant over time; They would cancel out in the two-way fixed effect model [@huntington-klein]. The standard errors were clustered at the classroom level.

In addition, I presented an unrestricted and adjusted DID model: $$\Delta score_{i,c} = \alpha + \beta \cdot migr_{i,c} + \gamma \cdot FE_{c} + \lambda \cdot score_{i,c;base} + \theta \cdot X_{i,c} + \varepsilon_{i,c}$$ where $X_{i,c}$ is the vector of covariates capturing the characteristics of children, their parents, and their households. The control variables were measured in the baseline survey. This version of model lifts the restriction that covariates from the baseline survey would be associated with a coefficient that equals one.

@@

The table below summarizes the changes in migration status in the sample. All 4040 children were living together with parents in one household in baseline, but about fifteen percent saw at least one parent migrated during the two waves of survey. In these new migrant households, only-mother migration and only-father migration each accounted for two out of five cases. The remaining one-fifth was both-parents migration.

| (1) No migration | (2) Any Parent | (3) Only Mother | (4) Mother (Uncond.) | (5) Only Father | (6) Father (Uncond.) | (7) Both Parents |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| 3415             | 625            | 253             | 381                  | 244             | 372                  | 128              |

: Parental migration status at follow-up survey. Among the columns: (2) = (3) + (5) + (7), (4) = (3) + (7), (6) = (5) + (7).

The following two tables summarizes the control variables. The distributions between girls and boys, urban and rural hukou holders, as well as Han Chinese and ethnic minorities largely align with the figures reported in the 2015 one-percent National Population Sample Survey [@nbs2017].

| Name          | Explanation                                      | Value             |    N |   \% |
|:--------------|---------------|:--------------|--------------:|--------------:|
| ch.girl       | gender                                           | 0 = Male          | 1980 | 49.0 |
|               |                                                  | 1 = Female        | 2060 | 51.0 |
| ch.rural      | hukou type                                       | 0 = Rural         | 2212 | 54.8 |
|               |                                                  | 1 = Urban         | 1828 | 45.2 |
| ch.health     | self-rated health                                | 1 = Very poor     |   21 |  0.5 |
|               |                                                  | 2 = Not very good |  112 |  2.8 |
|               |                                                  | 3 = Moderate      |  800 | 19.8 |
|               |                                                  | 4 = Good          | 1369 | 33.9 |
|               |                                                  | 5 = Very good     | 1738 | 43.0 |
| ch.boarding   | boarding at school                               | 0 = No            | 2885 | 71.4 |
|               |                                                  | 1 = Yes           | 1155 | 28.6 |
| ch.ethn.mnrt  | ethnic minority other than Han Chinese           | 0 = No            | 3758 | 93.0 |
|               |                                                  | 1 = Yes           |  282 |  7.0 |
| ch.skip.grd   | has skipped grade                                | 0 = No            | 3996 | 98.9 |
|               |                                                  | 1 = Yes           |   44 |  1.1 |
| ch.repeat.grd | has repeated grade                               | 0 = No            | 3662 | 90.6 |
|               |                                                  | 1 = Yes           |  378 |  9.4 |
| ch.preschool  | attended preschool                               | 0 = No            |  628 | 15.5 |
|               |                                                  | 1 = Yes           | 3412 | 84.5 |
| ch.aspr.univ  | aspires to finish university bachelor's or above | 0 = No            | 1158 | 28.7 |
|               |                                                  | 1 = Yes           | 2882 | 71.3 |

: The control variables at children's level

| Name          | Explanation                                                   | Value                         |    N |   \% |
|:--------------|---------------|:--------------|--------------:|--------------:|
| par.edu.year  | years of education of the best-educated parent                | 0 = no formal education       |    8 |  0.2 |
|               |                                                               | 6 = primary school            |  235 |  5.8 |
|               |                                                               | 9 = junior high school        | 1554 | 38.5 |
|               |                                                               | 12 = senior high school       | 1239 | 30.7 |
|               |                                                               | 15 = associate college        |  383 |  9.5 |
|               |                                                               | 16 = university undergraduate |  542 | 13.4 |
|               |                                                               | 19 = university postgraduate  |   79 |  2.0 |
| par.ccp       | at least one parent is Communist Party member                 | 0 = No                        | 3498 | 86.6 |
|               |                                                               | 1 = Yes                       |  542 | 13.4 |
| par.expt.univ | parent expects child to finish university bachelor's or above | 0                             |  704 | 17.4 |
|               |                                                               | 1                             | 3336 | 82.6 |
| home.internet | internet access at home                                       | 0                             | 1285 | 31.8 |
|               |                                                               | 1                             | 2755 | 68.2 |
| home.book     | number of extracurricular books at home, self-rated by child  | 1 = Very few                  |  383 |  9.5 |
|               |                                                               | 2 = Not many                  |  406 | 10.0 |
|               |                                                               | 3 = Some                      | 1298 | 32.1 |
|               |                                                               | 4 = Quite a few               | 1103 | 27.3 |
|               |                                                               | 5 = A great number            |  850 | 21.0 |
| home.finances | household financial situation self-rated by parent            | 1 = Very poor                 |  113 |  2.8 |
|               |                                                               | 2 = Somewhat poor             |  583 | 14.4 |
|               |                                                               | 3 = Moderate                  | 3102 | 76.8 |
|               |                                                               | 4 = Somewhat rich             |  230 |  5.7 |
|               |                                                               | 5 = Very rich                 |   12 |  0.3 |

: The control variables at parent's and household level

## Multiple Regression
